---
title: "Dry_Tortugas_Temp"
author: "Maddy Enda"
date: 01/26/2026
format: html
editor: visual
---

# Temperature Data 2026 Update: Dry Tortugas National Park

### Author- Madison Enda, January 26th, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   In this script, we will be combining all "master list" temperature files (in .xlsx format) off of the FWRI Coral Research Program's shared drive, and creating an index to be used as our primary key for the 'Temperature' table in the upcoming database. For ease of access (and to avoid syncing with SharePoint which proved to take some time),the folder was downloaded manually to my device, and thus my file path is unique to my device.

-   This workflow will include general cleaning (standardizing naming conventions, removing superfluous data, etc.) and the final result will be uploaded to the shared drive once again, until permissions are received to migrate to a new platform.

## Import Libraries

```{r}
# Install the readxl package to convert from .xlsx files. Comment out once done to avoid re-running
# install.packages("readxl")

# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, readxl, purrr)
```


## Loading in the .xlsx files

```{r}
# Grab file path where master files are
folder_path <- "C:/Users/Madison.enda/Desktop/Temp_Data/temp_dry_tortugas_2026_update"

# List all .xlsx files
file_list <- list.files(path = folder_path, pattern = "\\.xlsx$", full.names = TRUE)
```

## Read the files into a single data frame

```{r}
# Use the purrr package to apply read_excel to all files
df_list <- purrr::map(file_list, read_excel)
```




