---
title: "Dry_Tortugas_Temp"
author: "Maddy Enda"
date: 01/26/2026
format: html
execute: 
  echo: false
  eval: false
  warning: false
editor: visual
---

# Temperature Data 2026 Update: Dry Tortugas National Park

### Author- Madison Enda, January 26th, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   In this script, we will be combining all "master list" temperature files (in .xlsx format) off of the FWRI Coral Research Program's shared drive, and creating an index to be used as our primary key for the 'Temperature' table in the upcoming database. For ease of access (and to avoid syncing with SharePoint which proved to take some time),the folder was downloaded manually to my device, and thus my file path is unique to my device.

-   This workflow will include general cleaning (standardizing naming conventions, removing superfluous data, etc.) and the final result will be uploaded to the shared drive once again, until permissions are received to migrate to a new platform.

## Import Libraries

```{r}
# Install the readxl package to convert from .xlsx files. Comment out once done to avoid re-running
# install.packages("readxl")

# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, readxl, purrr, hms)
```

## Loading in the .xlsx files

```{r}
# Grab file path where master files are
folder_path <- "C:/Users/Madison.enda/Desktop/Temp_Data/temp_dry_tortugas_2026_update"

# List all .xlsx files
file_list <- list.files(path = folder_path, pattern = "\\.xlsx$", full.names = TRUE)
```

## Read the files into a single data frame

```{r}
# Use the purrr package to apply read_excel to all files
df_list <- suppressWarnings(purrr::map(file_list, read_excel)) # I suppressed warnings as it warned me that every date did not come in date format lol

# Combine using bind_rows since same columns exist
combined_temp <- bind_rows(df_list)
```

## View the combined data frame

```{r}
# Get some stats about our new data frame:

# Number of columns and number of rows
cat("Rows:", nrow(combined_temp), "Columns:", ncol(combined_temp), "\n")

# Take a look at the bottom of the df
tail(combined_temp)
```

## Ensure all columns have data in correct type and format

```{r}
# Use str to inspect the column name, type, and format
str(combined_temp)
```

-   Here we can see that site_id, Date, and Time need to be converted before we move forward, because: 1.) Date is currently in date + time format, but there is not time recorded in this column 2.) Time is simply a number, and being recorded in decimal foramt which uses more memory 3.) siteid is being recorded as a decimal, which uses more memory

## Converting to proper date format

```{r}
# Use as.Date on the Date column to convert 
combined_temp$Date <- as.Date(combined_temp$Date)

# Specify format of date as "yyyy-mm-dd", or 2006-09-15
format(combined_temp$Date,"%Y-%m-%d")
```

## Converting to hours, mins, secs in military time

```{r}
# Convert to integer before applying time
combined_temp$Time <- as.integer(combined_temp$Time)

# Convert the integer to a hour
combined_temp$Time <- hms::hms(hours = combined_temp$Time)
```

## Converting site_id to integer

```{r}
# Set the siteid column as an integer
combined_temp$siteid <- as.integer(combined_temp$siteid)
```

## Check one last time to see if type changes worked

```{r}
# Use str to inspect the column name, type, and format
str(combined_temp)
```

## Ensure we have no missing data or other such issues

```{r}
# Find the sum of all NA values in each column
colSums(is.na(combined_temp))
```

```{r}
# Check to see if the highest and lowest temps make sense
max(combined_temp$`Temp F`)
min(combined_temp$`Temp F`)
```

-   It seems we have three NA values in Temp F for some reason, so I will find them and fill them with the appropriate value converting from Fahrenheit

-   Secondly, we seem to have temperature values that are much higher than would be expected.I will filter to all observations over 101 (the record high recorded in 2023) and see how many there are, as well as values below 65 (from what I can research, the average minimum recorded historical value).

## Find which rows have the NA values

```{r}
# Identify the missing rows and confirm if Temp C has data
missing_temp_F <- combined_temp %>%
  filter(is.na(`Temp F`))

# Since Temp C has data for all three of these observations (all Palmata Patch in 2016 a couple hours apart), we will use case_when to replace the NA values with the correct temp in Fahrenheit converting from the Celsius value
combined_temp <- combined_temp %>%
  mutate(
    `Temp F` = case_when(
      is.na(`Temp F`) ~ `Temp C` * 1.8 + 32,
      TRUE            ~ `Temp F`
    )
  )

# Check again to see if we have any NA values
colSums(is.na(combined_temp))
```

## Find erroneous temperature values

```{r}
# Take the rows where the Temp F column has values higher than 101 degrees
high_temp_rows <- combined_temp %>%
  filter(`Temp F` > 100)

# Take the rows where the Temp F column has values lower than 68 degrees
low_temp_rows <- combined_temp %>%
  filter(`Temp F` < 65)
```

-   After checking in with other members of the team, we found the cold anomalies to correspond with the 2009-2011 cold shocks that did dramatically affect reefs. Seeing as how the two observations from 2018 occurred at 3:00 AM and 4:00 AM in January, I will be leaving those values as well.

-   Looking at all the high temp anomalies, they all seemed to occur in September of 2018, and I found that all values exceeding 95 degrees came from the same location at the same time. Therefore, after manually checking the master list from 2018, I have determined that all values above 92 degrees Fahrenheit (all from Texas Rock in 2018, from the same 24 hour time period)

```{r}
# Remove extraneous values
filtered_temp <- combined_temp %>%
  filter(`Temp F` < 92 )
```

## Assigning a primary key: observation ID

```{r}
# Create a column called Observation_ID
filtered_temp <- filtered_temp %>%
  arrange(Date) %>% # oldest → newest
  mutate(Observation_ID = row_number()) # oldest value gets ID = 1
```

## Cleaning Names & Exporting

```{r}
# Rename old columns to be snake case (`Temp C` to Temp_C)  
dry_tortugas_master_2026 <- filtered_temp %>%
  rename(
    observation_id = Observation_ID,
    date = Date,
    time = Time,
    site_name = sitename,
    site_id = siteid,
    temp_C = `Temp C`,
    temp_F = `Temp F`
  )
```

-   I now have a cleaned data frame of temperature from 2006-2021, we can export this. Normally, I would export directly to a server or database, but we unfortunately do not have this set up at the moment, so it will be saved in my current working directory, where the files were drawn from in the beginning.

## Write the final csv

```{r}
# Write the final data frame to a csv in my current working directory
write.csv(dry_tortugas_master_2025, "dry_tortugas_master_2025.csv", row.names = FALSE)
```

# Wrap Up:

I will be adding other master csv files to this "dry_tortugas_master_2025.csv", until we have it updated with all our data through 2025. However, while some years have master csv files (2019-2021), some do not have master lists available on the same drive, and therefore I will have to build those out before adding them (2022-2025).
