---
title: "Dry_Tortugas_Temp_2022-2025"
author: "Maddy Enda"
format: html
editor: visual
---

# Temperature Data 2026 Update: Dry Tortugas National Park

### Author- Madison Enda, January 27th, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   In this script, I will be looking into the Dry Tortugas temperature data collected from 2022-2025. After sifting through the data sets, I quickly realized that incongruities between them required them to be separated out into different .qmd files. Thus, the beginning of this script shows some investigation and the latter is the separate cleaning of the 2023 and 2022 csv files.

-   This data was pulled from the FWRI Coral Research Program's shared temperature data drive, and contain temperature data from the individual sites for each year, recorded once every hour.The final master file will be added to the new 2026 database. For ease of access (and to avoid syncing with SharePoint which proved to take some time),the folders were downloaded manually to my device, and thus my file path is unique to my device.

## Import Libraries

```{r}
# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, purrr, hms, lubridate)
```

## Load in the csv files for each year 2022-2023, creating master files to investigate

-   We will be loading the files in as individual years first, to see if they have the same columns, before attempting a join of all years

```{r}
# Using my local path, unpack all the files and read them in as a single csv per year
dt_temp_2022 <- list.files(
  path = "C:/Users/Madison.enda/Desktop/Temp_Data/trimmed", # The 2022 file was called "trimmed" in drive
  pattern = "\\.csv$",
  full.names = TRUE
) %>%
  map_dfr(read_csv)

dt_temp_2023 <- list.files(
  path = "C:/Users/Madison.enda/Desktop/Temp_Data/2023",
  pattern = "\\.csv$",
  full.names = TRUE
) %>%
  map_dfr(read_csv)
```

-   It unfortunately appears as though these files have many more columns than the old data. However, after running head() on the data sets, it seems like most of them have NA values in these columns. I will perform some basic analysis to check

## Simple Investigation

```{r}
# Checking the column names in 2022 (which has 1 more than the others)
colnames(dt_temp_2022)
```

-   It looks like a single row may have been added as other columns? I'm not quite sure, but I'll keep looking

```{r}
# Checking the column names of each set individually
colnames(dt_temp_2023)
```

-   This one is extremely confusing. It seems as though each "title" was a column, and all the data simply saved into the first column for each observation? Not quite sure how this occurred.

-   Since the 2023 data is in such a strange format, I'm going to load in one csv to check what each site looks like.

```{r}
# Specify file path of one csv file 
folder_path <- "C:/Users/Madison.enda/Desktop/Temp_Data/2023"

# Read CSVs by combining the folder path and file name
test_csv <- read_csv(file.path(folder_path, "BirdKey_Off1_22-23.csv"))
```

-   It looks like each file has all of the data condensed into one column, but not all data is separated by commas or semicolons. I will have to format them as I load them into a list, and then combine them into a single df. I will likely have to do additional cleaning afterwards as a result of the lack of delimiters.

## 2022 Data:

### Add Row and Remove Extra Columns

The 2022 data has one row that was accidentally recorded as separate columns, so this information needs to be deleted and a new row added for this data

```{r}
# Remove extra rows
dt_temp_2022 <- dt_temp_2022 %>%
  select(-c(`10/10/2021`,`11`, `Loggerhead Patch`, `46`, `29.615`, `85.307`))

# Rename the rows to match naming conventions
dt_temp_2022 <- dt_temp_2022 %>%
  rename(
    date = Date,
    time = Time,
    site_name = Sitename,
    site_id = Siteid,
    temp_C = `Temp C`,
    temp_F = `Temp F`
  )

# Add an observation_id column
dt_temp_2022 <- dt_temp_2022 %>%
  arrange(date) %>% # oldest → newest
  mutate(observation_id = row_number()) # oldest value gets ID = 1
```

### Check for NAs

```{r}
# Find the sum of all NA values in each column
colSums(is.na(dt_temp_2022))

# Take a look at the NAs
dt_2022_NAs <- dt_temp_2022 %>%
  filter(is.na(temp_C))
```

-   These columns are just all NA all across. There is no evidence of date's or where it came from, so they must be removed.

```{r}
# Remove all rows that had NAs in them
dt_temp_2022 <- dt_temp_2022 %>%
  filter(!is.na(temp_C))

# Check to see if there are NAs left somehow
colSums(is.na(dt_temp_2022))
```

### Begin formatting the 2022 data in accordance with the 2006-2021 process

```{r}
# Check data types for formatting
str(dt_temp_2022)

# Change the date to date time:

# Use as.Date on the Date column to convert 
dt_temp_2022$date <- as.Date(dt_temp_2022$date, format = "%m/%d/%Y")

# Specify format of date as "yyyy-mm-dd", or 2006-09-15
format(dt_temp_2022$date,"%Y-%m-%d")

# Change the time to military time:

# Convert to integer before applying time
dt_temp_2022$time <- as.integer(dt_temp_2022$time)

# Convert the integer to a hour
dt_temp_2022$time <- hms::hms(hours = dt_temp_2022$time)

# Set site_id as integer

# Set the site_id column as an integer
dt_temp_2022$site_id <- as.integer(dt_temp_2022$site_id)
```

### Final 2022 check

```{r}
# Check data types for formatting
str(dt_temp_2022)
```



_____________________________________________________________________________________________________



## 2023 Data:

### Begin formatting the 2023 data in accordance with the 2006-2021 process

```{r}
# Attempt to clean the 2023 file into 6 columns

# Function to fix a single malformed CSV
fix_csv <- function(file_path, n_cols = 6, sep = ",") {

  df <- read_csv(
    file_path,
    col_names = FALSE,
    show_col_types = FALSE
  )

  names(df) <- "raw_data"

  df %>%
    separate(
      col = raw_data,
      into = paste0("col", 1:n_cols),
      sep = sep,
      fill = "right",
      extra = "drop",
      convert = TRUE
    ) %>%
    mutate(source_file = basename(file_path))
}

# Get all CSV files in the folder
files <- list.files(
  path = "C:/Users/Madison.enda/Desktop/Temp_Data/2023",
  pattern = "\\.csv$",
  full.names = TRUE
)

# Fix each CSV, then combine into one data frame
dt_temp_2023 <- map_dfr(files, fix_csv)
```


### Check NA's and format

```{r}
# Look for NAs in any column
colSums(is.na(dt_temp_2022))
```
-   We need a Site Name to add to this data, so I need to see how many sites are in the source_file column to determine which site names need to replace the abbreviations

```{r}
# Use unique() to determine the different file names
unique(dt_temp_2023$source_file)
```

-   Ok, this strategy at least separates our observation_id, date-time, and temperature. All I need to do now is delete some rows, rename the columns, and separate some data

-   To ensure that I am using the correct column names, I will import the dry_tortugas_master_2025.csv and look at the unique values there as well

```{r}
# Import the dry_tortugas_master_2025.csv
dry_tortugas_master_2025 <- read_csv("dry_tortugas_master_2025.csv")

# Look at the unique values in that column
unique(dry_tortugas_master_2025$site_name)
```

### Add Site Names

-   There seems to be quite a few typos, which we will fix once all years have been added together, but for now we are gonna stick to correct spellings, and use that for our case_when() to change the file names into site names

```{r}
# Remove the first two rows
dt_temp_2023 <- dt_temp_2023 %>% 
  slice(-(1:2))

# Use case_when to change the csv file names in source_file into site_name
dt_temp_2023 <- dt_temp_2023 %>%
mutate(site_name = case_when(
    source_file == "BCR_Off1_22-23.csv"        ~ "Black Coral Rock",
    source_file == "BirdKey_Off1_22-23.csv"    ~ "Bird Key Reef",
    source_file == "Davis_Off1_22-23.csv"      ~ "Davis Rock",
    source_file == "Loggerhead_Off1_22-23.csv" ~ "Loggerhead Patch",
    source_file == "Mayers_Off2_22-23.csv"     ~ "Mayer's Peak",
    source_file == "Maze_Off1_22-23.csv"       ~ "The Maze",
    source_file == "Palmata_Off2_22-23.csv"    ~ "Palmata Patch",
    source_file == "Prolifera_In2_22-23.csv"   ~ "Prolifera",
    source_file == "Temptation_Off2_22-23.csv" ~ "Temptation",
    source_file == "Texas_In2_22-23.csv"       ~ "Texas Rock",
    source_file == "WhiteShoal_Off3_22-23.csv" ~ "White Shoal",
    TRUE                                       ~ NA_character_  # fallback for any other file names
  ))

# Drop extra columns
dt_temp_2023 <- dt_temp_2023 %>% 
  select(-c(col4, col5, col6, source_file))

# Rename columns
dt_temp_2023 <- dt_temp_2023 %>%
  rename(
    observation_id = col1,
    date = col2,
    temp_F = col3
  )
```

-   Alright, now we have some of the data we need in the right columns. I will separate the time columns, and check for NAs.

```{r}
# We will separate the dates out into date and time
dt_temp_2023 <- dt_temp_2023 %>%
  separate(date, into = c("date", "time"), sep = " ")
```
### Remove erroneous data

```{r}
# Find the sum of all NA values in each column
colSums(is.na(dt_temp_2023))

# We will remove all 10 rows that have no temperature data (NA or empty)
dt_temp_2023 <- dt_temp_2023 %>%
  filter(!is.na(temp_F) & temp_F != "")
```
-   Great! Now we don't have any NA or missing values. Let's check now for any other strange inputs by using grepl() on our data set. Normally, we would use unique(), but seeing as how observation_id is all unique, it would give us over 97,000 unique things. So instead we will look for all values that do not contain only numbers

```{r}
# Look for erraneous values in our data set
dt_2023_errors <- dt_temp_2023 %>%
  filter(grepl("[^0-9]", observation_id))

# Check out the non-numbers
head(dt_2023_errors)
```

-   This shows us that we have 10 rows with \# as the input, and these need to be removed as well

```{r}
# Remove the # observations in the observation_id column (10 rows)
dt_temp_2023 <- dt_temp_2023 %>%
  filter(!grepl("[^0-9]", observation_id))
```

-   Lastly, before our final check, we will add a temp_C columns

```{r}
# Create the temperature column in Celsius
dt_temp_2023 <- dt_temp_2023 %>%
  mutate(temp_F = as.numeric(temp_F),
         temp_C = (temp_F - 32) / 1.8)
```

## Check to ensure cleanliness

```{r}
# Find the sum of all NA values in each column
colSums(is.na(dt_temp_2023))
```

## Begin formatting the 2023 data in accordance with the 2006-2021 process

```{r}
# Formatting the data set:

# Check data types for formatting
str(dt_temp_2023)

# Change the date to date time:

# Use mdy on the Date column to convert 
dt_temp_2023$date <- mdy(dt_temp_2023$date)

# Specify format of date as "yyyy-mm-dd", or 2006-09-15
format(dt_temp_2023$date,"%Y-%m-%d")

# Change the time to military time:

# Convert the integer to a hour
dt_temp_2023$time <- as_hms(dt_temp_2023$time)


# Set the site_id column as an integer
dt_temp_2023$observation_id <- as.integer(dt_temp_2023$observation_id)
```

______________________________________________________________________________________________________

## Save cleaned master lists to master list folder on my local computer

```{r}
# Write the final data frame to a csv in my current working directory
write.csv(dt_temp_2022, "C:/Users/Madison.Enda/Desktop/Temp_Data/temp_dry_tortugas_2026_update/DRTO_2022_Master.csv")

write.csv(dt_temp_2023, "C:/Users/Madison.Enda/Desktop/Temp_Data/temp_dry_tortugas_2026_update/DRTO_2023_Master.csv")
```


