---
title: "Dry_Tortugas_Temp_2025"
author: "Maddy Enda"
format: html
editor: visual
---

# Temperature Data 2026 Update: Dry Tortugas National Park 2025

### Author- Madison Enda, January 28th, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   In this script, we will be combining the site specific temperature data sets from 2025 into master lists, aggregating them into a combined csv, and appending that to the dry_tortugas_master_2025.csv.

-   This data was pulled from the FWRI Coral Research Program's shared temperature data drive, and contain temperature data from the individual sites for each year, recorded once every hour.The final master file will be added to the new 2026 database. For ease of access (and to avoid syncing with SharePoint which proved to take some time),the folders were downloaded manually to my device, and thus my file path is unique to my device.

## Load libraries

```{r}
# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, purrr, hms, lubridate)
```


## Use function and code from previous files (Dry_Tortugas_Temp_2022-2023.qmd) to load and clean data

```{r}
# Fix the 2025 data

# Function to fix a single malformed CSV
fix_csv <- function(file_path, n_cols = 6, sep = ",") {

  df <- read_csv(
    file_path,
    col_names = FALSE,
    show_col_types = FALSE
  )

  names(df) <- "raw_data"

  df %>%
    separate(
      col = raw_data,
      into = paste0("col", 1:n_cols),
      sep = sep,
      fill = "right",
      extra = "drop",
      convert = TRUE
    ) %>%
    mutate(source_file = basename(file_path))
}

# Get all CSV files in the folder
files <- list.files(
  path = "C:/Users/Madison.enda/Desktop/Temp_Data/DT",
  pattern = "\\.csv$",
  full.names = TRUE
)

# Fix each CSV, then combine into one data frame
dt_temp_2025 <- map_dfr(files, fix_csv)

# Remove the first two rows
dt_temp_2025 <- dt_temp_2025 %>% 
  slice(-(1:2))

# Use case_when to change the csv file names in source_file into site_name
dt_temp_2025 <- dt_temp_2025 %>%
mutate(site_name = case_when(
    source_file == "BCR_Off1_24-25.csv"        ~ "Black Coral Rock",
    source_file == "BirdKey_Off1_24-25.csv"    ~ "Bird Key Reef",
    source_file == "Davis_Off1_24-25.csv"      ~ "Davis Rock",
    source_file == "Loggerhead_Off1_24-25.csv" ~ "Loggerhead Patch",
    source_file == "Mayers_Off2_24-25.csv"     ~ "Mayer's Peak",
    source_file == "Maze_Off1_24-25.csv"       ~ "The Maze",
    source_file == "Palmata_Off2_24-25.csv"    ~ "Palmata Patch",
    source_file == "Prolifera_In2_24-25.csv"   ~ "Prolifera",
    source_file == "Temptation_Off2_24-25.csv" ~ "Temptation",
    source_file == "Texas_In2_24-25.csv"       ~ "Texas Rock",
    source_file == "WhiteShoal_Off3_24-25.csv" ~ "White Shoal",
    TRUE                                       ~ NA_character_  # fallback for any other file names
  ))

# Drop extra columns
dt_temp_2025 <- dt_temp_2025 %>% 
  select(-c(col4, col5, col6, source_file))

# Rename columns
dt_temp_2025 <- dt_temp_2025 %>%
  rename(
    observation_id = col1,
    date = col2,
    temp_F = col3
  )

# We will separate the dates out into date and time
dt_temp_2025 <- dt_temp_2025 %>%
  separate(date, into = c("date", "time"), sep = " ")

# We will remove all 10 rows that have no temperature data (NA or empty)
dt_temp_2025 <- dt_temp_2025 %>%
  filter(!is.na(temp_F) & temp_F != "")

# Remove the # observations in the observation_id column (10 rows)
dt_temp_2025 <- dt_temp_2025 %>%
  filter(!grepl("[^0-9]", observation_id))

# Create the temperature column in Celsius
dt_temp_2025 <- dt_temp_2025 %>%
  mutate(temp_F = as.numeric(temp_F),
         temp_C = (temp_F - 32) / 1.8)
```

