---
title: "Investigating_Dry_Tortugas_Temp"
author: "Maddy Enda"
format: html
editor: visual
---

# Investigating the Dry Tortugas Temperature Data- 2025 Master List

### Author: Maddy Enda

### Written: January 29th, 2026

## Overview:

-   The 2025 Dry Tortugas Temperature data set was compiled in January of 2026 as a part of a database management project. The data, collected from HoboWare autonomous sensors placed on each survey site in the Dry Tortugas, was aggregated from the FWRI Coral Research Program's One Drive.

-   In this script, we will be looking through the data, performing basic summary statistics and visualizations to ensure data was collected and cleaned correctly before continuing with other data sets.

## Import Libraries

```{r}
# Use the librarian package to import many libraries at once'
librarian::shelf(tidyverse,
                 dplyr,
                 here,
                 hms,
                 purrr,
                 ggplot2,
                 gt,
                 lubridate)
```

## Load in the dry tortugas temperature file

```{r}
# Import the master file temperature csv for 2025
dry_tortugas_master_2025 <- read_csv("dry_tortugas_master_2025.csv")
```

## Investigate data

```{r}
# Create summary statistics using summarise()
 summary_table <- dry_tortugas_master_2025%>%
   mutate(year = lubridate::year(date)) %>%
  group_by(year) %>%
  summarise(
    mean_temp   = mean(temp_F, na.rm = TRUE),
    median_temp = median(temp_F, na.rm = TRUE),
    sd_temp     = sd(temp_F, na.rm = TRUE),
    min_temp    = min(temp_F, na.rm = TRUE),
    max_temp    = max(temp_F, na.rm = TRUE),
    n_obs       = n()
  )

# Display table using
summary_table %>%
  gt() %>%
  fmt_number(
    columns = -year,
    decimals = 2
  ) %>%
   tab_options(
    table.font.color = "black",
    data_row.padding = px(12),
    column_labels.padding = px(12)
  ) %>%
  tab_header(
    title = "Dry Tortugas Annual Temperature Summary",
    subtitle = "Temperature in Degrees Fahrenheit by Year"
  )
```

-   Ok, so after checking with the original scripts, its looking like 2022-2025 are the only ones where observation counts could be amiss. I will look into how
