---
title: "Florida_Keys_Temp_2001-2021"
author: "Maddy Enda"
format: html
editor: visual
---

# Temperature Data 2026 Update: Florida Keys National Marine Sanctuary

### Author- Madison Enda, February 2nd, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   In this script, we will be combining all "master list" temperature files (in .xlsx format) off of the FWRI Coral Research Program's shared drive, and creating an index to be used as our primary key for the 'Temperature' table in the upcoming database. For ease of access (and to avoid syncing with SharePoint which proved to take some time),the folder was downloaded manually to my device, and thus my file path is unique to my device.

-   This workflow will include general cleaning (standardizing naming conventions, removing superfluous data, etc.) and the final result will be uploaded to the shared drive once again, until permissions are received to migrate to a new platform.

## Import Libraries

```{r}
# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, readxl, purrr, hms, lubridate)
```

## Loading in the .csv files

```{r}
# Grab file path where master files are
folder_path <- "C:/Users/Madison.enda/Desktop/Temp_Data/FKNMS_2002-2021"

# List all .xlsx files
file_list <- list.files(path = folder_path, pattern = "\\.csv$", full.names = TRUE)
```

## Read the files into a single data frame

```{r}
# Use the purrr package to apply read_excel to all files
df_list <- suppressWarnings(purrr::map(file_list, read_csv)) # I suppressed warnings as it warned me that every date did not come in date format lol

# Combine using bind_rows since same columns exist
combined_temp <- bind_rows(df_list)
```

## View the combined data frame

```{r}
# Get some stats about our new data frame:

# Number of columns and number of rows
cat("Rows:", nrow(combined_temp), "Columns:", ncol(combined_temp), "\n")

# Take a look at the bottom of the df
tail(combined_temp)
```

## Check rows with NA values

```{r}
# Check to see if the empty ID rows can be removed
colSums(is.na(combined_temp))
```

-   It looks like the ID columns 7-79 can all be removed, since all rows are empty

```{r}
# Remove empty columns
combined_temp <- combined_temp %>%
  select(Date, Time, SiteName, SiteID, TempC, TempF)
```

-   Now let's find where the other NA values are in the data

```{r}
# Filter to rows where Date is NA
no_dates <- combined_temp %>%
  filter(is.na(Date))

# Ensure that there is no data where the Date is also NA
colSums(is.na(no_dates))

# Remove these rows from the data set
combined_temp <- combined_temp %>%
  filter(!is.na(Date))
```

-   Now, let's check any rows that may have NA values and see what needs to be corrected before data types can be changed across the board

```{r}
# Keep rows where any column has NA
rows_with_na <- combined_temp %>%
  filter(if_any(everything(), is.na))
```

-   It looks like only two rows of the remaining rows with NAs actually contain valuable information. It seems as though they somehow got put into the incorrect format, but they do contain data. Thus I will manually add the information contained in these rows:

5/6/2003 16 Tennessee Shallow 55 27.554

5/13/2013 15 Tennessee Deep 75 26.475

and remove all rows that remain with any NAs in them since they will not contain any unique data.

```{r}
# Remove rows where any column has NA
combined_temp <- combined_temp %>%
  filter(if_all(everything(), ~ !is.na(.)))

# Now add the two rows of data that were in the wrong format
combined_temp <- combined_temp %>%
  add_row(Date = "05/06/2003",
          Time = 16,
          SiteName = "Tennessee Shallow",
          SiteID = 55,
          TempC = 27.554,
          TempF = 81.597)

# Check to see if the above worked before proceeding

# Add the second row
combined_temp <- combined_temp %>%
    add_row(Date = "05/13/2013",
          Time = 15,
          SiteName = "Tennessee Deep",
          SiteID = 75,
          TempC = 26.475,
          TempF = 79.655)
```

## Ensure all columns have data in correct type and format

```{r}
# Use str to inspect the column name, type, and format
str(combined_temp)
```

-   Here we can see that site_id, Date, and Time need to be converted before we move forward, because: 1.) Date is currently in date + time format, but there is not time recorded in this column 2.) Time is simply a number, and being recorded in decimal foramt which uses more memory 3.) siteid is being recorded as a decimal, which uses more memory

## Converting to proper date format

```{r}
# Find any erroneous dates by identifying odd patterns. We don't want to use unique() or distinct() here because that will return millions of characters lol
date_patterns <- combined_temp %>%
  mutate(
    date_pattern = gsub("[0-9]", "D", Date)
  ) %>%
  count(date_pattern, sort = TRUE)
```

-   There are 8 distinct patterns of date saved here, I will use their structure to convert them all to what we are looking for with the lubridate() package

```{r}
# Use lubridate::parse_date_time() function to get dates
combined_temp$Date <- parse_date_time(
  combined_temp$Date,
  orders = c("mdy", "mdy HM", "mdy HMS")
)
```

## Converting to hours, mins, secs in military time

```{r}
# Convert to integer before applying time
combined_temp$Time <- as.integer(combined_temp$Time)

# Convert the integer to a hour
combined_temp$Time <- hms::hms(hours = combined_temp$Time)
```

## Converting site_id to integer

```{r}
# Set the siteid column as an integer
combined_temp$SiteID <- as.integer(combined_temp$SiteID)
```

## Check one last time to see if type changes worked

```{r}
# Use str to inspect the column name, type, and format
str(combined_temp)
```

## Ensure we have no missing data or other such issues

```{r}
# Find the sum of all NA values in each column
colSums(is.na(combined_temp))
```

## Change column names to fit current conventions

```{r}
# Change column names to fit conventions
combined_temp <- combined_temp %>%
  rename(
    date = Date,
    time = Time,
    site_name = SiteName,
    site_id = SiteID,
    temp_C = TempC,
    temp_F = TempF
  )
```

```{r}
# Check to see if the highest and lowest temps make sense
max(combined_temp$temp_F)
min(combined_temp$temp_F)
```

-   We seem to have temperature values that are much higher than would be expected.I will filter to all observations over 101 (the record high recorded in 2023) and see how many there are, as well as values below 65 (from what I can research, the average minimum recorded historical value).

## Find erroneous temperature values

```{r}
# Take the rows where the Temp F column has values higher than 92 degrees
high_temp_rows <- combined_temp %>%
  filter(temp_F > 95)

# Take the rows where the Temp F column has values lower than 68 degrees
low_temp_rows <- combined_temp %>%
  filter(temp_F < 59)
```

-   After checking in with other members of the team, we found the cold anomalies to correspond with the 2009-2011 cold shocks that did dramatically affect reefs. Seeing as how the two observations from 2018 occurred at 3:00 AM and 4:00 AM in January, I will be leaving those values as well.

-   Looking at all the high temp anomalies, they all seemed to occur in September of 2018, and I found that all values exceeding 95 degrees came from the same location at the same time. Therefore, after manually checking the master list from 2018, I have determined that all values above 92 degrees Fahrenheit (all from Texas Rock in 2018, from the same 24 hour time period)

```{r}
# Remove extraneous values
filtered_temp <- combined_temp %>%
  filter(temp_F < 95 )
```

## Assigning a primary key: observation ID

```{r}
# Create a column called Observation_ID
florida_keys_master_2025 <- filtered_temp %>%
  arrange(date) %>% # oldest → newest
  mutate(observation_id = row_number()) # oldest value gets ID = 1
```

-   I now have a cleaned data frame of temperature from 2006-2021, we can export this. Normally, I would export directly to a server or database, but we unfortunately do not have this set up at the moment, so it will be saved in my current working directory, where the files were drawn from in the beginning.

## Write the final csv

```{r}
# Write the final data frame to a csv in my current working directory
write.csv(florida_keys_master_2025, "florida_keys_master_2025.csv", row.names = FALSE)
```

# Wrap Up:

I will be adding other master csv files to this "florida_keys_master_2025.csv", until we have it updated with all our data through 2025. However,the years 2022, 2023, 2024, and 2025 do not have master files in place, so they will have to be created and cleaned before I can join. In the next script, I will be working with the 2022 data.
