---
title: "Florida_Keys_Temp_2022"
author: "Maddy Enda"
format: html
editor: visual
---

# Temperature Data 2026 Update: Florida Keys National Marine Sanctuary

### Author- Madison Enda, February 5th, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   In this script, we will be combining all 2022 temperature files from the Florida Key National Marine Sancturay: pulling them off of the FWRI Coral Research Program's shared drive and creating an index to be used as our primary key for the 'Temperature' table in the upcoming database. For ease of access (and to avoid syncing with SharePoint which proved to take some time),the folder was downloaded manually to my device, and thus my file path is unique to my device.

-   This workflow will include general cleaning (standardizing naming conventions, removing superfluous data, etc.) and the final result will be uploaded to the shared drive once again, until permissions are received to migrate to a new platform.

## Import Libraries

```{r}
# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, purrr, hms, lubridate)
```

## Loading in the .xlsx files

```{r}
# Grab file path where master files are
folder_path <- "C:/Users/Madison.enda/Desktop/Temp_Data/2022"

# List all .csv files
file_list <- list.files(path = folder_path, pattern = "\\.csv$", full.names = TRUE)
```

## Read the files into a single data frame

```{r}
# Use the purrr package to apply read_excel to all files
df_list <- suppressWarnings(purrr::map(file_list, read_csv)) # I suppressed warnings as it warned me that every date did not come in date format lol

# Combine using bind_rows since same columns exist
combined_temp <- bind_rows(df_list)
```

## View the combined data frame

```{r}
# Get some stats about our new data frame:

# Number of columns and number of rows
cat("Rows:", nrow(combined_temp), "Columns:", ncol(combined_temp), "\n")

# Take a look at the bottom of the df
tail(combined_temp)
```

# Check to see if empty additional columns have any data

```{r}
# Look for the sum of all NA values in each row
colSums(is.na(combined_temp))
```

-   Ok, so there are some NA values in the core 6 columns, but also some actual results in the additional columns. Let's separate them out and combine them again once they are clean

```{r}
# Select the added columns
extra_columns <- combined_temp %>%
  select(Date, Time, SiteName, SiteID, TempC, TempF, `...7`, `...8`) 

# Select only relevant columns
combined_temp <- combined_temp %>%
  select(Date, Time, Sitename, Siteid, `Temp C`, `Temp F`)
```

-   Now, we can analyze them separately

```{r}
# Remove columns ...7 and ...8
extra_columns <- extra_columns %>%
  select(Date, Time, SiteName, SiteID, TempC, TempF)

# Find columns without NA values for the temperatures (either C or F)
extra_columns_clean <- extra_columns %>%
  filter(!is.na(TempC) | !is.na(TempF))

# Use extra_columns_clean as our new df for the observations that will be added

# Check the same thing for the main dataset
combined_temp_clean <- combined_temp %>%
  filter(!is.na(`Temp C`) | !is.na(`Temp F`))
```

-   Let's rename the columns of the data sets, so we can bind them!

```{r}
# Rename the columns of both data sets
combined_temp <- combined_temp_clean %>%
  rename(
    "date" = "Date",
    "time" = "Time",
    "site_name" = "Sitename",
    "site_id" = "Siteid",
    "temp_C" = "Temp C",
    "temp_F" = "Temp F"
  )

# Rename the columns that will be added to the main data
extra_columns_clean <- extra_columns_clean %>%
  rename(
    "date" = "Date",
    "time" = "Time",
    "site_name" = "SiteName",
    "site_id" = "SiteID",
    "temp_C" = "TempC",
    "temp_F" = "TempF" 
  )

# Combine the two 2022 temp data frames
fk_temp_2022 <- bind_rows(combined_temp, extra_columns_clean)
```

-   Now we check to see which NA values are left over...

```{r}
# Look for the sum of all NA values in each row
colSums(is.na(fk_temp_2022))
```

## Converting data types:

-   Nice! There are no NA values in our data set now. It's time to convert data types

```{r}
# Look at data types in the fk_temp_2022 data 
str(fk_temp_2022)
```

### Converting dates using lubridate()

```{r}
# Find any erroneous dates by identifying odd patterns. We don't want to use unique() or distinct() here because that will return millions of characters lol
date_patterns <- combined_temp %>%
  mutate(
    date_pattern = gsub("[0-9]", "D", date)
  ) %>%
  count(date_pattern, sort = TRUE)
```

-   There are 4 distinct patterns of date saved here, I will use their structure to convert them all to what we are looking for with the lubridate() package

```{r}
# Use lubridate::parse_date_time() function to get dates
fk_temp_2022$date <- parse_date_time(
  fk_temp_2022$date,
  orders = c("mdy", "mdy HM", "mdy HMS")
)
```

### Converting to hours, mins, secs in military time

```{r}
# Convert to integer before applying time
fk_temp_2022$time <- as.integer(fk_temp_2022$time)

# Convert the integer to a hour
fk_temp_2022$time <- hms::hms(hours = fk_temp_2022$time)
```

### Converting site_id to integer

```{r}
# Set the siteid column as an integer
fk_temp_2022$site_id <- as.integer(fk_temp_2022$site_id)
```

## Removing erroneous temperature values

```{r}
# Take the rows where the Temp F column has values higher than 92 degrees
high_temp_rows <- fk_temp_2022 %>%
  filter(temp_F > 90)

# Take the rows where the Temp F column has values lower than 68 degrees
low_temp_rows <- fk_temp_2022 %>%
  filter(temp_F < 68)
```

-   It appears as though all temperature values fall within the expected ranges for the Florida Keys. These csv files must have already been cleaned, which is great news! Now we can add an observation_id column and join the 2022 temp data to the 2002-2021 data.

## Assigning a primary key: observation_id

```{r}
# Create a column called Observation_ID
fk_temp_2022 <- fk_temp_2022 %>%
  arrange(date) %>% # oldest → newest
  mutate(observation_id = row_number()) # oldest value gets ID = 1

# Import the master temp data file
florida_keys_master_2025 <- read_csv("Florida_Keys_National_Marine_Sanctuary/florida_keys_master_2025.csv")

# Bind the rows together
florida_keys_master_2025 <- bind_rows(florida_keys_master_2025, fk_temp_2022)
```

## Write the new csv file to current working directory

```{r}
# Write the final data frame to a csv in my current working directory
write.csv(florida_keys_master_2025, "florida_keys_master_2025.csv", row.names = FALSE)
```
