---
title: "Master_Florida_Keys_Temp_2025"
author: "Maddy Enda"
format: html
editor: visual
---

# Temperature Data 2026 Update: Florida Keys National Marine Sanctuary Master File Creation

### Author- Madison Enda, February 9th, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   In this script, we create the final florida_keys_master_2025.csv file, with cleaned and updated information from 2002-2025. The data was all pulled from the FWRI Coral Research Program's sharepoint drive, and was joined and standardized in the other scripts in this repository. Therefore, we will be joining the 2023, 2024, and 2025 master files to the florida_keys_master_2025.csv, checking for temperature anomalies, and then exporting to this device, and then the organization's drive until such a time as we have permissions to upload it to a server.

## Load libraries

```{r}
# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, purrr, hms, stringr)
```

-   First, let's load in the large master file for all years: florida_keys_master_2025.csv

```{r}
# Import the master file from this working directory
florida_keys_master_2025 <- read_csv("florida_keys_master_2025.csv")
```

## Combine the 2023, 2024, and 2025 data

```{r}
# Import the 2023 file
fk_temp_2023 <- read_csv("C:/Users/Madison.enda/Desktop/Temp_Data/Temperature/Florida_Keys_National_Marine_Sanctuary/florida_keys_temp_2023.csv")

# Import the 2024 file
fk_temp_2024 <- read_csv("C:/Users/Madison.enda/Desktop/Temp_Data/Temperature/Florida_Keys_National_Marine_Sanctuary/florida_keys_temp_2024.csv")

# Import the 2025 file
fk_temp_2025 <- read_csv("C:/Users/Madison.enda/Desktop/Temp_Data/Temperature/Florida_Keys_National_Marine_Sanctuary/florida_keys_temp_2025.csv")

# Combine the 2024 and 2025 data sets
fk_temp_2023_2025 <- bind_rows(fk_temp_2023, fk_temp_2024, fk_temp_2025)
```

## Check for structural integrity and NAs

```{r}
# Find the sum of all NA values in each column
colSums(is.na(fk_temp_2023_2025))
```

## Add a site_id column and populate with NAs to fill later with a case_when()

```{r}
# Create a site_id column that is just NAs
fk_temp_2023_2025 <- fk_temp_2023_2025 %>%
  mutate(site_id = NA)
```

# Bring together the 2023-2025 data and the master list

-   It's finally time to bring it all together! We will bind the master list to the 2024 and 2025 data, then fill the missing values in site_id (should be 207,545) with the corresponding values using site_name. Then we recreate the observation_id column, run through our final checks, and it's done!

```{r}
# Use bind_rows to join the data
florida_keys_master_2025 <- bind_rows(florida_keys_master_2025, fk_temp_2023_2025)
```

-   Now, let's take a look at our structure and types

```{r}
# Find the sum of all NA values in each column
colSums(is.na(florida_keys_master_2025))

# Look at column names
colnames(florida_keys_master_2025)

# Identify data types
str(florida_keys_master_2025)
```

## Fill the NA values in site ID

-   First, we will identify all relevant site names, and correct any misspellings

```{r}
sort(unique(florida_keys_master_2025$site_name))
```

```{r}
# Replace misspellings
florida_keys_master_2025$site_name <- str_replace_all(florida_keys_master_2025$site_name, c(
  "West Washer Women" = "West Washerwoman",
  "Jaap Reef Reef" = "Jaap Reef",
  "Port Patch" = "Porter Patch",
  "Pamlata Patch" = "Palmata Patch",
  "Sobrero Shallow" = "Sombrero Shallow",
  "Red Dunn" = "Red Dun Reef",
  "BurrFish" = "Burr Fish",
  "Smit Shoal" = "Smith Shoal"
))
```

-   Now that we have all sites, we need to fill the empty NA in site_id with the corresponding value for each site_name

```{r}
# Group by site_name and fill with site_id value
florida_keys_master_2025 <- florida_keys_master_2025 %>%
  group_by(site_name) %>%
  mutate(site_id = coalesce(site_id, first(site_id[!is.na(site_id)]))) %>%
  ungroup()
```

-   Now let's check for any remaining NA values in any column

```{r}
# Find the sum of all NA values in each column
colSums(is.na(florida_keys_master_2025))

# Look at column names
colnames(florida_keys_master_2025)

# Identify data types
str(florida_keys_master_2025)
```

## Reassign character types and create new observtion_id

```{r}
# Set the site_id column as an integer
florida_keys_master_2025$site_id <- as.integer(florida_keys_master_2025$site_id)

# Set the observation_id column as an integer
florida_keys_master_2025$observation_id <- as.integer(florida_keys_master_2025$observation_id)

# Recount the observation_id column
florida_keys_master_2025 <- florida_keys_master_2025 %>%
  arrange(date) %>% # oldest → newest
  mutate(observation_id = row_number()) %>% # oldest value gets id of 1
  select(observation_id, everything()) # moves observation_id to be the first column
```

## Final Cleaning for the 2025 Dry Tortugas Temperature Data

```{r}
# Look at the minimum and maximum temperatures
max(florida_keys_master_2025$temp_F)
min(florida_keys_master_2025$temp_F)
```

## Save the final file!

```{r}
# Write our final file to my current working directory
write.csv(florida_keys_master_2025, "florida_keys_master_2025.csv", row.names = FALSE)
```
