---
title: "Southeastern_Florida_Temp_2018"
author: "Maddy Enda"
format: html
editor: visual
---

# Temperature Data 2026 Update: Southeastern Florida

### Author- Madison Enda, February 11th, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   In this script, we will be cleaning the SECREMP 2018 temperature data, and compiled from individual site temperature files.

## Import Libraries

```{r}
# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, readxl, purrr, hms, lubridate)
```

## Loading in the .csv files

```{r}
secremp_2017_2018 <- read_csv("C:/Users/Madison.Enda/Desktop/Temp_Data/SECREMP_Master_Files/temp_2018-2022/SECREMP_1718_Temperature.csv")
```

The columns that we are looking to make in order to be standardized with the CREMP data are: - date - time - temp_C - temp_F - site - site_id - region - habitat - station - observation_id

First, let's convert our Year, Month, and Day columns into a date column!

### Creating a date column

```{r}
# Make a date column from Year, Month, and Day
secremp_2017_2018$date <- make_date(secremp_2017_2018$Year, secremp_2017_2018$Month, secremp_2017_2018$Day)

# Remove the Year, Month, Day columns
secremp_2017_2018 <- secremp_2017_2018 %>%
  select(-c(Year, Month, Day))
```

### Make the time column an hour in military format

```{r}
# Convert to integer before applying time
secremp_2017_2018$Time <- as.integer(secremp_2017_2018$Time)

# Convert the integer to a hour
secremp_2017_2018$Time <- hms::hms(hours = secremp_2017_2018$Time)
```

### Rename the columns

```{r}
# Rename old columns to be snake case (`Temp C` to Temp_C)  
secremp_2017_2018 <- secremp_2017_2018 %>%
  rename(
    time = Time,
    site = Site,
    temp_C = TempC,
    temp_F = TempF,
    station_id = StationID
  )
```

### Assigning a primary key: observation_id

```{r}
# Create a column called Observation_ID
secremp_2017_2018 <- secremp_2017_2018 %>%
  arrange(date) %>% # oldest → newest
  mutate(observation_id = row_number()) # oldest value gets ID = 1

# Reorder the columns in the dataframe
secremp_2017_2018 <- secremp_2017_2018 %>%
  select(observation_id, date, everything())
```

```{r}
colSums(is.na(secremp_2017_2018))
```

### Filter out erroneous data:

There appear to be three observations where the temperature in Celsius is missing and for some reason Fahrenheit is 32 degrees. This seems very wrong, so I will be removing these values.

The highest temperature of 95 degrees was recorded a couple of hours before that logger is recorded as being picked up, in the afternoon/early evening of June 28th, 2017. Therefore that value will be kept in.

```{r}
# Remove the odd freezing values
secremp_2017_2018 <- secremp_2017_2018 %>%
  filter(temp_F > 32)
```

### Write the file to our current working directory'

```{r}
# Write the final data frame to a csv in my current working directory
write.csv(secremp_2017_2018, "southeastern_florida_temp_2017_2018.csv", row.names = FALSE)
```
