---
title: "Southeastern_Florida_Temp_2020"
author: "Maddy Enda"
format: html
editor: visual
---

# Temperature Data 2026 Update: Southeastern Florida

### Author- Madison Enda, February 11th, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   In this script, we will be cleaning the SECREMP 2020 temperature data, and compiled from individual site temperature files.

## Import Libraries

```{r}
# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, readxl, purrr, hms, lubridate)
```

## Loading in the .csv files

```{r}
secremp_temp_2020 <- read_csv(
  "C:/Users/Madison.Enda/Desktop/Temp_Data/SECREMP_Master_Files/temp_2018-2022/SETEMP_2020.csv",
  locale = locale(encoding = "Latin1"),
  col_types = cols(Time = col_character())
)
```

The columns that we are looking to make in order to be standardized with the CREMP data are: - date - time - temp_C - temp_F - site - site_id - region - habitat - station - observation_id

First, let's convert our Year, Month, and Day columns into a date column!

### Convert the date column to date data types

```{r}
# Make the date column a date object
secremp_temp_2020$Date <- mdy(secremp_temp_2020$Date)
```

### Convert time column to hms object

```{r}
# Convert from character strings to time
secremp_temp_2020$Time <- as.hms(secremp_temp_2020$Time)
```

### Create a temp_F column

```{r}
# Convert the temp_C column to Fahrenheit
secremp_temp_2020 <- secremp_temp_2020 %>%
  mutate(temp_F = `Temp °C` * 1.8 + 32)
```

### Change column names for standardization

```{r}
# Rename old columns to be snake case (`Temp C` to Temp_C)  
secremp_temp_2020 <- secremp_temp_2020 %>%
  rename(
    date = Date,
    time = Time,
    temp_C = `Temp °C`,
    station_id = StationID
  )

# Create a column called Observation_ID
secremp_temp_2020 <- secremp_temp_2020 %>%
  arrange(date) %>% # oldest → newest
  mutate(observation_id = row_number()) # oldest value gets ID = 1

# Reorder the columns in the dataframe
secremp_temp_2020 <- secremp_temp_2020 %>%
  select(observation_id, date, time, temp_C, temp_F, station_id)
```

```{r}
# Check for any NA values
colSums(is.na(secremp_temp_2020))
```

There are a couple NA values for temperature; we will need to remove them before we proceed.

```{r}
# Filter out the NA values 
secremp_temp_2020 <- secremp_temp_2020 %>%
  filter(!is.na(temp_C))
```

### Write the file to our current working directory'

```{r}
# Write the final data frame to a csv in my current working directory
write.csv(secremp_temp_2020, "southeastern_florida_temp_2019_2020.csv", row.names = FALSE)
```
