---
title: "Southeastern_Florida_Temp_Deliverable"
author: "Maddy Enda"
format: html
editor: visual
---

# Temperature Data 2026 Update: Southeastern Florida

### Author- Madison Enda, February 12th, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   This script cleans and looks-over the previous SECREMP cumulative temperature deliverable. This data set has records from 2007-2024. Added to this master list will be the 2025 data.

## Import Libraries

```{r}
# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, readxl, purrr, hms, lubridate)
```

## Loading in the .csv files

```{r}
secremp_temp_2007_2024 <- read_csv("C:/Users/Madison.Enda/Desktop/SECREMP_Temperature_2007-2024.csv")
```

The columns that we are looking to make in order to be standardized with the CREMP data are: - date - time - temp_C - temp_F - site - site_id - region - habitat - station - observation_id

First, let's convert our Year, Month, and Day columns into a date column!

### Creating a date column

```{r}
# Make a date column from Year, Month, and Day
secremp_temp_2007_2024$date <- make_date(secremp_temp_2007_2024$Year, secremp_temp_2007_2024$Month, secremp_temp_2007_2024$Day)

# Remove the Year, Month, Day columns
secremp_temp_2007_2024 <- secremp_temp_2007_2024 %>%
  select(-c(Year, Month, Day))
```

### Make the time column an hour in military format

```{r}
# Convert to integer before applying time
secremp_temp_2007_2024$Time <- as.integer(secremp_temp_2007_2024$Time)

# Convert the integer to a hour
secremp_temp_2007_2024$Time <- hms::hms(hours = secremp_temp_2007_2024$Time)
```

### Rename the columns

```{r}
# Rename old columns to be snake case (`Temp C` to Temp_C)  
secremp_temp_2007_2024 <- secremp_temp_2007_2024 %>%
  rename(
    time = Time,
    temp_C = TempC,
    temp_F = TempF,
    station_id = StationID,
    site_id = SiteID,
    site_name = SiteName
  )
```

### Assigning a primary key: observation_id

```{r}
# Create a column called Observation_ID
secremp_temp_2007_2024 <- secremp_temp_2007_2024 %>%
  arrange(date) %>% # oldest → newest
  mutate(observation_id = row_number()) # oldest value gets ID = 1

# Reorder the columns in the dataframe
secremp_temp_2007_2024 <- secremp_temp_2007_2024 %>%
  select(observation_id, date, time, everything())
```

```{r}
# There appears to be one observation where the temperature in Celsius is missing which is odd, considering we calculate Fahrenheit from Celcius typically.
no_C <- secremp_temp_2007_2024 %>%
filter(is.na(secremp_temp_2007_2024$temp_C))

# The associated observation is extremely cold, so I believe that we should remove this observation
# Let's check Broward county 5 in July of 2024. It is possible that the value in Fahrenheit is meant to be recorded in Celsius, but I want to be sure before changing it.
july_broward_county_5 <- secremp_temp_2007_2024 %>%
  filter(site_name == "Broward County 5" & date == '2024-07-27')
```

### Filter out erroneous data:

The temperature around 2024-07-27 did not get up to 32 degrees C, or about 80 degrees Fahrenheit, according to the surrounding values. A quick logger check shows that there were was a gap in recording during that time. This observation will be removed

```{r}
# Remove the odd freezing values
secremp_temp_2007_2024 <- secremp_temp_2007_2024 %>%
  filter(temp_F > 32)
```

### Add any extra rows that may be requried (Habitat)

```{r}
# Create df with only habitat and site_id
joining_df <- southeastern_florida_2007_2017 %>%
  select(site_id, habitat) %>%
  distinct()

# Remove the typo at site 91
joining_df <- joining_df %>%
  filter(habitat != "First Reef")

# Left join the habitat column with the southeastern_florida_master_2025
updated_secremp_master <- southeastern_florida_master_2025 %>%
  left_join(
  joining_df %>%
    select(site_id, habitat),
  by = "site_id")
```

### Write the final file

```{r}
# Write the final data frame to a csv in my current working directory
write.csv(updated_secremp_master, "southeastern_florida_master_2025.csv", row.names = FALSE)
```

