---
title: "Southeastern_Florida_Temp_2022"
author: "Maddy Enda"
format: html
editor: visual
---

# Temperature Data 2026 Update: Southeastern Florida

### Author- Madison Enda, February 12th, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   In this script, we will be cleaning the SECREMP 2022 temperature data, and compiled from individual site temperature files.

## Import Libraries

```{r}
# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, readxl, purrr, hms, lubridate)
```

## Loading in the .csv files

```{r}
secremp_temp_2022 <- read_csv(
  "C:/Users/Madison.Enda/Desktop/Temp_Data/SECREMP_Master_Files/temp_2018-2022/SETEMP_2022.csv",
  locale = locale(encoding = "Latin1"),
  col_types = cols(Time = col_character())
)
```

### Check data types

```{r}
# Look at our data types using str()
str(secremp_temp_2022)
```

### Convert the date column to date data types

```{r}
# Make the date column a date object
secremp_temp_2022$Date <- mdy(secremp_temp_2022$Date)
```

### Find strange date time formats

```{r}
# Take a look at the SECREMP date patterns for 2019
date_patterns <- secremp_temp_2022 %>%
  mutate(
    date_pattern = gsub("[0-9]", "D", Date)
  ) %>%
  count(date_pattern, sort = TRUE)
```

-   It appears as though we have NA values for some reason. Let's take a look at them

```{r}
# Look at the NA values in date
NA_values <- secremp_temp_2022 %>%
  filter(is.na(Date))

# Remove the NA values as they contained no data
secremp_temp_2022 <- secremp_temp_2022 %>%
  filter(!is.na(Date))
```

-   After the NA values were removed, the data type changed to Date, so we do not need to convert this column again

### Convert time column to hms object

```{r}
# Convert from character strings to time
secremp_temp_2022$Time <- as.hms(secremp_temp_2022$Time)
```

### Create a temp_F column

```{r}
# Convert the temp_C column to Fahrenheit
secremp_temp_2022 <- secremp_temp_2022 %>%
  mutate(temp_F = `Temp °C` * 1.8 + 32)
```

### Change column names for standardization

```{r}
# Rename old columns to be snake case (`Temp C` to Temp_C)  
secremp_temp_2022 <- secremp_temp_2022 %>%
  rename(
    date = Date,
    time = Time,
    site_code = `Site Code`,
    temp_C = `Temp °C`,
    station_id = stationid
  )

# Create a column called Observation_ID
secremp_temp_2022 <- secremp_temp_2022 %>%
  arrange(date) %>% # oldest → newest
  mutate(observation_id = row_number()) # oldest value gets ID = 1

# Reorder the columns in the dataframe
secremp_temp_2022 <- secremp_temp_2022 %>%
  select(observation_id, date, time, temp_C, temp_F, everything())
```

```{r}
colSums(is.na(secremp_temp_2022))
```

-   There doesn't appear to be any extremely high or low temperature observations, so we will leave it all for now! :)

### Write the file to our current working directory'

```{r}
# Write the final data frame to a csv in my current working directory
write.csv(secremp_temp_2022, "csvs/southeastern_florida_temp_2021_2022.csv", row.names = FALSE)
```
