---
title: "Southeastern_Florida_Temp_2021"
author: "Maddy Enda"
format: html
editor: visual
---

# Temperature Data 2026 Update: Southeastern Florida

### Author- Madison Enda, February 11th, 2026

### Contact- madison.enda\@myfwc.com

## Overview:

-   In this script, we will be cleaning the SECREMP 2021 temperature data, and compiled from individual site temperature files.

## Import Libraries

```{r}
# Install librarian for multiple packages and quick readability
# install.packages("librarian")

# Using librarian::shelf to import many libraries at once
librarian::shelf(tidyverse, dplyr, here, readxl, purrr, hms, lubridate)
```

## Loading in the .csv files

```{r}
secremp_temp_2021 <- read_csv(
  "C:/Users/Madison.Enda/Desktop/Temp_Data/SECREMP_Master_Files/temp_2018-2022/SETEMP_2021.csv")
```

### Cleaning the data set

```{r}
# Remove the extra columns
secremp_temp_2021 <- secremp_temp_2021 %>%
  select(-c(Year, Month, Day, Sort, Time...7, ))
```

### Check data types

```{r}
# Look at our data types using str()
str(secremp_temp_2021)
```

### Convert the date column to date data types

```{r}
# Make the date column a date object
secremp_temp_2021$Date <- mdy(secremp_temp_2021$Date)
```


### Change column names for standardization

```{r}
# Rename old columns to be snake case (`Temp C` to Temp_C)  
secremp_temp_2021 <- secremp_temp_2021 %>%
  rename(
    date = Date,
    time = Time...6,
    temp_C = TempC,
    temp_F = TempF,
    site = Site,
    station_id = StationID
  )

# Create a column called Observation_ID
secremp_temp_2021 <- secremp_temp_2021 %>%
  arrange(date) %>% # oldest → newest
  mutate(observation_id = row_number()) # oldest value gets ID = 1

# Reorder the columns in the dataframe
secremp_temp_2021 <- secremp_temp_2021 %>%
  select(observation_id, date, time, temp_C, temp_F, site, station_id)
```

```{r}
# Check for any NA values
colSums(is.na(secremp_temp_2021))
```

### Write the file to our current working directory'

```{r}
# Write the final data frame to a csv in my current working directory
write.csv(secremp_temp_2021, "southeastern_florida_temp_2020_2021.csv", row.names = FALSE)
```


